---
title: "Explortory Anlaysis"
subtitle: "EDA"
output: github_document
---

Setting options for plots & embedding

```{r, include = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continous.colour = "viridis",
  ggplot2.continous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Load the weather data

```{r, include = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>% #the lubridate package, a tidyverse adjacent package, which take the "floor" of the date and pulls it to the nearest month (so all dates in Jan-xx-2001, will be noted as Jan 01 - 01)
  select(name, id, everything())

weather_df
```


## `group_by`

```{r}
weather_df %>% 
  group_by(name, month) %>% #3 names & 12 months, so should have around 36 groups
  ungroup() #undoing the grouping you did ... you can also ungroup one variable at a time ungroup(name)
```


## counting things

count month observations
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```


we can use `count` as well (kinda synonymous with group_by if you're using group_by + summarize() just for counting)

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```


**Never** use base R's `table` function.
Reason:
1 - outputs are not in a dataframe (it's a vector) 
2 - to get to the `table` function, we had to `pull` a variable out of a df 
```{r, eval = F}
weather_df %>% 
  pull(month) %>% 
  table()
```



Other helpful counters

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(), 
    n_days = n_distinct(date) #n_distinct here is looking at all the dates in a given month (what  you grouped by)
    )
```


## A digression on 2x2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold", 
      tmax >= 5 ~ "not cold", 
      TRUE  ~ ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n()) 
```


using `janitor::tabyl()` to make a 2x2 table given group_by + summarize-d output

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold", 
      tmax >= 5 ~ "not cold", 
      TRUE  ~ ""
    )
  ) %>% 
  janitor::tabyl(name, cold) #making a 2x2 the easy way (or whatever by whatever table)
```


## General summaries

computing in the different months, the average tmax ... and a couple more summaries 
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = T), #first define your variable name (mean_tmax) and then define what you want to go into that summary variable
    mean_prcp = mean(prcp, na.rm = T),
    median_tmin = median(tmin, na.rm = T)
  )
```


This is a dataframe! SO I'm going to make graphs off of it
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = T),
    mean_prcp = mean(prcp, na.rm = T),
    median_tmin = median(tmin, na.rm = T)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() + 
  geom_line()

#compare this with a non-summarized plot version of the same data: 

weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_line()
```


The point ^ from this above chunk is that you can do group_by then summarize and then make a plot off of it


**Now, Suppose you want to summarize many columns.**
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean)) #computing the mean of all the variables starting from the prcp variables TO [:] tmin
```


Reminder: sometimes your results are easier to read in another format. 

*Peep* the `knitr::kable` function, which outputs a nice table output when you knit your Rmd to an md
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1)
```


## `group_by` and `mutate`
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax #a centered t_max ... which is tmax - mean of tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
  geom_point() # here each data point is centered around zero

#note how the weather_df is still the same size ... you're just adding the same mean_tmax for every obs in a given month (which is what you grouped by)
```



## Window Functions

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    temp_rank = min_rank (tmax)) %>% #ranking the temperatures according to their max temperatures
    #temp_rank = min_rank (desc(tmax))  #"descending" tmax
    filter(temp_rank == 1) #this filter by gives us the coldest day in each of the 3 places, by month
```


`lag`


```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    temp_change = tmax - lag(tmax)) %>% 
    summarize(
      temp_change_max = max(temp_change, na.rm = TRUE),
      temp_change_sd = sd(temp_change, na.rm = TRUE)
    )
##largest temperture change in a day by place

```


