---
title: "Visualization"
subtitle: "Part I"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)

```


```{r, include = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots
Create a scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

```

Name approach, same plot - using the " %>% " pipe function
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Save and edit plot object. 

```{r}
#saving my ggplot "frame" into an object I can then manipulate later on
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

#manipulating the ggplot object I created
weather_plot + geom_point()
```

## Advanced scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


## what about `aes` placement .. ?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + 
  geom_smooth(se = T)
```

Let's facet some things! 


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) + 
  geom_point(alpha = .2) + 
  geom_smooth(se = T) + 
  facet_grid(. ~ name) 
```

Let's combine some elements and try a new plot.

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) + 
  geom_smooth(se = F) + 
  facet_grid(. ~ name)
```


## Some small notes 

How many geoms do you need? You can have whatever geoms you want
 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = T)
```

You can use a neat geom!

```{r}
#hexagons
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) + 
  geom_hex() 
  
#squares
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) + 
  geom_bin2d()

#contours
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) + 
  geom_density2d() + 
  geom_point(alpha = .3)
```
 
 
## Univariate Plots
Where you have just one variable

HISTOGRAMS

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) + 
  geom_histogram() 
```


Can we add color to the ^ histogram ... ?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) + 
  geom_histogram(position = "dodge")

#color vs fill in aes()
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram(position = "dodge")

#faceting as opposed to having all 3 distbs in one plot
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram() + 
  facet_grid(. ~ name)


```

In the above histograms, it's quite hard to compare one histogram with another. 
So, we're going to use a different geom to be able to compare multiple histograms of multiple variables

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density(alpha = .3)

#with `adjust`
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density(alpha = .3, adjust = .5)

```

On `density` in `geom_density`
You can think of density in `geom_density` as a histogram that has had it individual points smoothed out. 
An argument FOR geom_density is that it makes it easier to compare histograms. 
But note how you lost some information compared to geom_histogram, which give you a more detailed description of your histogram. You can try and address by passing the `adjsut = ` argument in the geom_density funciton.


## Box Plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) + 
  geom_boxplot()
```


## Trendy plots
Other univariate plots

1st up - violin plots (nvm they don't quite look like violins)
argument for `geom_violin` is tha they convey some more information, as a graphic, on the distribution. Taht's to say, it's easier to tell whether something is left skewed/right skewed/symmterical. 
Think of violin plots as a cross between histograms and density plots. 
Also note how violin plots are symmterical around the y-axis ... they're really just histograms that mirror eachother. 
Argument against violin plot in contrast to box_plots is that they don't give you a direct info on, say, the median. But you can address this by adding whaterver statistics you need using `stat_summary(fun = "median")`

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin(alpha = .5) + 
  stat_summary(fun = "median")
```


Trendy plots part II - Ridge Plots (most popular plot of 2017)


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name, fill = name)) + 
  geom_density_ridges()
```


## SAVE & EMBED

Let's save a scatterplot. 

```{r}
weather_plot = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

weather_plot

ggsave("./plots/weather_plot.jpg", weather_plot, width = 8 , height = 5)
```

What about embeddig ... ? i.e., can we control what a figure will end up looking like in your knitted md? 

```{r}
weather_plot
```


embedding using chunk options (at different size)
```{r, fig.width=12, fig.height=6}
weather_plot
```

```{r, fig.width=6, fig.height=3}
weather_plot
```





